---
title: "p8105_hw2_nk2814"
author: "Naama Kipperman"
date: "9/26/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1
### Read and clean Mr. Trash Wheel data sheet

```{r}

# First, load necessary packages.

library(tidyverse)
library(readxl)
library(dplyr)

# Import excel file and name it 'p1_trash', specify sheet
# Omit non-data entries (rows with notes / figures; columns containing notes)
# Clean up variable names
# Omit rows where dumpster is missing
# Round the number of sports balls to the nearest integer and converts the result to an integer variable (using as.integer)

p1_trash=read_excel(path="./data/healthyharbor.xlsx", sheet=1, col_names = TRUE, range=cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate( 
      sports_balls =  as.integer(round(sports_balls, digits=0)))
 
```

### Read and clean precipitation data for 2017
```{r}

# Read in 2017 sheet
# Clean variable names
# Omit rows without precipitation data
# Add a variable year
p1_2017 = read_excel(path="./data/healthyharbor.xlsx", sheet=4, col_names = TRUE, skip=1, range="A2:B14") %>% 
    janitor::clean_names() %>%
  rename(precipitation=total) %>% 
  drop_na(precipitation) %>% 
  mutate(
    year=2017
  )


```

### Read and clean precipitation data for 2018

```{r}

# Read in 2018 sheet
# Clean variable names
# Omit rows without precipitation data
# Add a variable year (2018)
p1_2018 = read_excel(path="./data/healthyharbor.xlsx", sheet=3, col_names = TRUE, skip=1, range="A2:B14") %>% 
    janitor::clean_names() %>%
  rename(precipitation=total) %>% 
  drop_na(precipitation) %>% 
  mutate(
    year=2018
  )



```


### Combine precipitation datasets and convert month to a character variable

```{r}
p1_combined = 
  full_join(p1_2017, p1_2018) %>% 
  rename(
    month.name=month
  )

```

### Write a paragraph about this data.


The number of observations in the 2017 precipitation dataset is 
`r nrow(p1_2017)`. The number of observations in the 2018 precipitation dataset is `r nrow(p1_2018)`. The primary variables in the datasets were `r names(p1_2017)`. The total precipitation in 2018 was `r sum(pull(p1_2018,precipitation)) ` inches. The median number of sports balls in a dumpster in 2017 was `r filter(p1_trash, year == 2017) %>% pull(sports_balls) %>% median() ` balls.

## Problem 2

### 1) Clean the data in pols-month.csv

```{r}

# Import csv file, clean names, and separate 'mon' into 3 integer variables year, month, day

politicians_data = read_csv(file = "./data/pols-month.csv") %>%
  janitor::clean_names() %>% 
  separate(mon, into=c("year", "month", "day"), sep="-") %>% 
  mutate( 
        year =  as.integer(year),
        month = as.integer(month),
        day = as.integer(day))
# Create a president variable taking values gop and dem, and remove prez_dem, prez_gop,day variable

politicians_data %>% 
mutate(presidents = if_else(prez_gop==1, 'gop', 'dem')) %>%
  select(-prez_dem, -prez_gop, -day)
  


```

First, clean the data in pols-month.csv. Use separate() to break up the variable mon into integer variables year, month, and day; replace month number with month name; create a president variable taking values gop and dem, and remove prez_dem and prez_gop; and remove the day variable.

create a president variable taking values gop and dem, and remove prez_dem and prez_gop; and remove the day variable.

clean the data in snp.csv
